\bf On semidefinite-representable sets over valuation fields

### 2025-05-16 article/paper_padic_spec.tex
algo formal
commit by Simone

\begin{algorithm}
    \caption{SOLVELP($A,b,c,D,e$)}\label{algo_LP}
    \begin{algorithmic}[1]
      \REQUIRE
      $A \in K^{d\times n}$, $b \in K^d$, $c \in K^n$, $D \in K^{m \times n}$ and $e \in K^m$
      \ENSURE [TYPE, $x$], where:
      \vspace{-0.25cm}
      \begin{itemize}
      \item[]
        TYPE $=$ INFEAS and $x=[]$ if and only if \eqref{LP} is infeasible
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ UNBOUND and $x=[]$ if and only if \eqref{LP} is feasible and unbounded
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ FEAS if and only if \eqref{LP} is feasible and has a solution $x$
      \end{itemize}
      \vspace{-0.1cm}
      \hrule
      \vspace{0.1cm}
      \IF{$Dx = e$ has no solution}
      \RETURN [INFEAS, []]
      \ENDIF
      \IF{$D \neq {\bf 0}$}
      \STATE find $J \in K^{n \times R}$, with $R = \dim\ker D$, such that $\text{Im}\, J = \ker D$
      \STATE find $x_0$ such that $Dx_0=e$
      \STATE $[\mathrm{STATUS}, y] \leftarrow \text{SOLVELP}(AJ,b+A x_0,J^T c,0,0)$
      \RETURN $[\mathrm{STATUS}, y]$
      \ENDIF
      %\FOR{epochs = $1$ to $N$}
      %\WHILE{$(j\le m)$}
      %\STATE Randomly initialize $w_i=\{w_1,w_2,\dots,w_n\}$
      %\STATE input $o_j=\{o_1,o_2,\dots,o_m\}$ in the input layer
      %\STATE forward propagate $(f_i\cdot w_i)$ through layers until getting the predicted result $y$
      %\STATE compute $e=y-y^2$
      %\STATE back propagate $e$ from right to left through layers
      %\STATE update $w_i$
      %\ENDWHILE
      %\ENDFOR
    \end{algorithmic}
\end{algorithm}

@@

\begin{proposition}\label{reducD}
  Let $A \in K^{d \times n}, b \in K^d, c\in K^n$, $D \in K^{m\times n}$,
  $e \in K^m$ and let $J \in K^{n \times R}$ be a matrix such that
  ${\rm Im}\, J = \ker D$ and $R = \dim \ker D$.
  Assume there exists $x_0 \in K^n$ such that $D x_0 = e$, and let
  $x \in \{z \in K^n : Dz = e\} = x_0 + \ker D$.
  Let $y \in K^R$ denote the unique element of $K^n$ such that
  $x = x_0 + J y$. Then:
  \begin{enumerate}
  \item\label{reducD_item1}
    $x$ is feasible for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is feasible for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$.
  \item\label{reducD_item2}
    $x$ is a solution for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is a solution for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$ and
    $\langle c,x^* \rangle = \langle J^T c,y^* \rangle$\footnote{\simone{Est-ce que
    c'est bien $\langle c,x^* \rangle = \langle J^T c,y^* \rangle$ ou plutôt la condition
    plus faible $\val \langle c,x^* \rangle = \val \langle J^T c,y^* \rangle$ ?
  Même question pour \Cref{solsLP}.}}.
  \end{enumerate}
\end{proposition}

\begin{proof}
  %The existence and uniqueness of $y$ are granted by the fact that $x \in x_0 + \ker D =x_0+ {\rm Im } J$  and $\ker J = \{0\}$. 
  First, if $D={\bf 0}$, then by the assumptions one must have $e=0$, and the problem
  is already reduced (with $J=I_n$ and $x_0=0$).
  Next, assume $D\neq{\bf 0}$, then $Dx=e$ and $Ax + b \le 0$ if and only if
  $AJy + b+Ax_0 \le 0$, proving \Cref{reducD_item1}. Finally, for all
  $c \in K^n$, $\langle c,x\rangle = \langle c, Jy\rangle = \langle J^T c,y\rangle$, which proves \Cref{reducD_item2}.
\end{proof}

@@

\Cref{solsLP} allows us to reduce \eqref{LP} to diagonal linear programs.

\begin{proposition} \label{solsLP}
  Let $A \in K^{d\times n}$, $b \in K^d$, $c\in K^n$ and let
  $S = \mathrm{SNF}(A) = QAP^{-1} \in K^{d\times n}$, with
  $P \in \GL_d(\OK)$ and $Q\in \GL_d(\OK)$.
  Define $b' = Qb$ and $c' = (P^{-1})^Tc$.
  Then
   \begin{enumerate}
   \item \label{solsLP_item1}
     $x \in K^n$ is feasible for \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y = P x \in K^n$ is
     feasible for \eqref{LP} with data $(S,b',c',0,0)$.
   \item \label{solsLP_item2}
     $x^* \in K^n$ is a solution to \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y^* = P x^*$ is
     a solution to \eqref{LP} with data $(S,b',c',0,0)$ and $\langle c,x^* \rangle = \langle c',y^* \rangle$.
   \end{enumerate}
\end{proposition}
\begin{proof}
  Note that the transition matrices $P$ and $Q$ defining SNF$(A)$ are invertible in $\OK$
  and as such they preserve inequalities: $Mz \geq 0$ if and only if $Nz \geq 0$, with
  $N = \mathrm{SNF}(M)$. Thus one has that $Ax+b = AP^{-1}y +b \geq 0$ if and only if
  $Q(AP^{-1}y + b) = Sy + b' \geq 0$; this yields \Cref{solsLP_item1}.
  Moreover, for all $c \in K^n$, we have $\langle c, x\rangle = \langle c, P^{-1}Px \rangle =
  \langle (P^{-1})^T c, Px \rangle = \langle c',y\rangle$, which proves
  \Cref{solsLP_item2}.
\end{proof}

\begin{proposition}\label{prop:reduc}
  Let $(A,b,c), (S,b',c') \in K^{d\times n} \times K^d \times K^n$ be as in \Cref{solsLP},
  where $S = \mathrm{SNF}(A)$. Then
  \begin{enumerate}
  \item \label{prop:reduc_it1}
    \eqref{LP} with data $(S,b',c',0,0)$ is feasible if and only if the $d-r$ last coefficients
    of $b'$ have nonnegative valuation, for $r = {\rm rank}\,S$.
  \item \label{prop:reduc_it2}
    Assume at least one of the $n-r$ last coefficients of $c'$ is non-zero. If \eqref{LP} with
    data $(S,b',c')$ is feasible, then it is unbounded and does not admit solutions.
  \end{enumerate}
\end{proposition}
\begin{proof}
  %%%%\corentin{préciser les notations}. 
  A vector $y = (y_1, \ldots, y_n)^T \in K^n$ is feasible for \eqref{LP} with data
  $(S,b',c',0)$ if and only if $\val(s_i y_i + b_i) \geq 0$ for\footnote{\simone{Ce n'est pas plutôt $\val(s_i y_i + b'_i) \geq 0$ ? Car c'est $b'$ et non pas $b$.}} all $i = 1,\ldots,n$.
  However $s_i = 0$ for $n-r \leq i\leq n$. Therefore, if \eqref{LP} with data $(S,b',c',0)$ is feasible,
  necessarily the $d-r$ last coefficients of $b'$ have nonnegative valuation.
  Conversely, if the $d-r$ last coefficients of $b'$ are zero, the vector $y \in K^n$ such that
  $y_i = -\frac{-b'_i}{s_i}$ if $1 \leq i \leq r$, and $y_i =0$ otherwise, is feasible for \eqref{LP}
  with data $(S,b',c',0,0)$.

>>>>

\begin{algorithm}
    \caption{\textbf{SOLVELP}($A,b,c,D,e$)}\label{algo_LP}
    \begin{algorithmic}[1]
      \REQUIRE
      $A \in K^{d\times n}$, $b \in K^d$, $c \in K^n$, $D \in K^{m \times n}$ and $e \in K^m$
      \ENSURE [TYPE, $x$], where:
      \vspace{-0.25cm}
      \begin{itemize}
      \item[]
        TYPE $=$ INFEAS and $x=[]$ if and only if \eqref{LP} is infeasible
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ UNBOUND and $x=[]$ if and only if \eqref{LP} is feasible and unbounded
      \vspace{-0.25cm}
      \item[]
        TYPE $=$ FEAS if and only if \eqref{LP} is feasible and has a solution $x$
      \end{itemize}
      \vspace{-0.1cm}
      \hrule
      \vspace{0.1cm}
      \IF{$Dx = e$ has no solution}
      \RETURN [INFEAS, []]
      \ENDIF
      \IF{$D \neq {\bf 0}$}
      \STATE find $J \in K^{n \times R}$, with $R = \dim\ker D$, such that $\text{Im}\, J = \ker D$
      \STATE find $x_0$ such that $Dx_0=e$
      \RETURN $\textbf{SOLVELP}(AJ,b+A x_0,J^T c,0,0)$
      \ENDIF
      \STATE $[Q,S,P] \leftarrow \textbf{SNF}(A)$
      \IF{$S \neq A$}
      \STATE $[\mathrm{STATUS}, y] \leftarrow \textbf{SOLVELP}(S,Qb,(P^{-1})^Tc,0,0)$
      \RETURN $[\mathrm{STATUS}, P^{-1}y]$
      \ELSE
      \STATE
      \ENDIF
    \end{algorithmic}
\end{algorithm}

@@

\begin{lemma}\label{reducD}
  Let $A \in K^{d \times n}, b \in K^d, c\in K^n$, $D \in K^{m\times n}$,
  $e \in K^m$ and let $J \in K^{n \times R}$ be a matrix such that
  ${\rm Im}\, J = \ker D$ and $R = \dim \ker D$.
  Assume there exists $x_0 \in K^n$ such that $D x_0 = e$, and let
  $x \in \{z \in K^n : Dz = e\} = x_0 + \ker D$.
  Let $y \in K^R$ denote the unique element of $K^n$ such that
  $x = x_0 + J y$. Then:
  \begin{enumerate}
  \item\label{reducD_item1}
    $x$ is feasible for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is feasible for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$.
  \item\label{reducD_item2}
    $x$ is a solution for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is a solution for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$ and
    $\langle c,x^* \rangle = \langle J^T c,y^* \rangle$\footnote{\simone{Est-ce que
    c'est bien $\langle c,x^* \rangle = \langle J^T c,y^* \rangle$ ou plutôt la condition
    plus faible $\val \langle c,x^* \rangle = \val \langle J^T c,y^* \rangle$ ?
  Même question pour \Cref{solsLP}.}}.
  \end{enumerate}
\end{lemma}
\begin{proof}
  %The existence and uniqueness of $y$ are granted by the fact that $x \in x_0 + \ker D =x_0+ {\rm Im } J$  and $\ker J = \{0\}$. 
  First, if $D={\bf 0}$, then by the assumptions one must have $e=0$, and the problem
  is already reduced (with $J=I_n$ and $x_0=0$).
  Next, assume $D\neq{\bf 0}$, then $Dx=e$ and $Ax + b \le 0$ if and only if
  $AJy + b+Ax_0 \le 0$, proving \Cref{reducD_item1}. Finally, for all
  $c \in K^n$, $\langle c,x\rangle = \langle c, Jy\rangle = \langle J^T c,y\rangle$, which proves \Cref{reducD_item2}.
\end{proof}

@@

\Cref{solsLP} allows us to reduce Problem \eqref{LP} without linear equations
to diagonal linear programs.

\begin{lemma} \label{solsLP}
  Let $A \in K^{d\times n}$, $b \in K^d$, $c\in K^n$ and let
  $S = \mathrm{SNF}(A) = QAP^{-1} \in K^{d\times n}$, with
  $P \in \GL_d(\OK)$ and $Q\in \GL_d(\OK)$.
  Define $b' = Qb$ and $c' = (P^{-1})^Tc$.
  Then
   \begin{enumerate}
   \item \label{solsLP_item1}
     $x \in K^n$ is feasible for \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y = P x \in K^n$ is
     feasible for \eqref{LP} with data $(S,b',c',0,0)$.
   \item \label{solsLP_item2}
     $x^* \in K^n$ is a solution to \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y^* = P x^*$ is
     a solution to \eqref{LP} with data $(S,b',c',0,0)$ and $\langle c,x^* \rangle = \langle c',y^* \rangle$.
   \end{enumerate}
\end{lemma}
\begin{proof}
  Note that the transition matrices $P$ and $Q$ defining SNF$(A)$ are invertible in $\OK$
  and as such they preserve inequalities: $Mz \geq 0$ if and only if $Nz \geq 0$, with
  $N = \mathrm{SNF}(M)$. Thus one has that $Ax+b = AP^{-1}y +b \geq 0$ if and only if
  $Q(AP^{-1}y + b) = Sy + b' \geq 0$; this yields \Cref{solsLP_item1}.
  Moreover, for all $c \in K^n$, we have $\langle c, x\rangle = \langle c, P^{-1}Px \rangle =
  \langle (P^{-1})^T c, Px \rangle = \langle c',y\rangle$, which proves
  \Cref{solsLP_item2}.
\end{proof}

Finally we show in \Cref{prop:reduc} that solving a diagonal \eqref{LP} amounts
to checking the valuations of input data:

\begin{lemma}\label{prop:reduc}
  Let $(A,b,c), (S,b',c') \in K^{d\times n} \times K^d \times K^n$ be as in \Cref{solsLP}.
  Then:
  \begin{enumerate}
  \item \label{prop:reduc_it1}
    \eqref{LP} with data $(S,b',c',0,0)$ is feasible if and only if the $d-r$ last coefficients
    of $b'$ have nonnegative valuation, for $r = {\rm rank}\,S$.
  \item \label{prop:reduc_it2}
    Assume at least one of the $n-r$ last coefficients of $c'$ is non-zero. If \eqref{LP} with
    data $(S,b',c')$ is feasible, then it is unbounded and does not admit solutions.
  \end{enumerate}
\end{lemma}
\begin{proof}
  %%%%\corentin{préciser les notations}. 
  A vector $y = (y_1, \ldots, y_n)^T \in K^n$ is feasible for \eqref{LP} with data
  $(S,b',c',0)$ if and only if $\val(s_i y_i + b_i) \geq 0$ for\footnote{\simone{Ce n'est pas plutôt $\val(s_i y_i + b'_i) \geq 0$ ? Car c'est $b'$ et non pas $b$.}} all $i = 1,\ldots,n$.
  However $s_i = 0$ for $n-r \leq i\leq n$. Therefore, if \eqref{LP} with data $(S,b',c',0)$ is feasible,
  necessarily the $d-r$ last coefficients of $b'$ have nonnegative valuation.
  Conversely, if the $d-r$ last coefficients of $b'$ are zero, the vector $y \in K^n$ such that
  $y_i = -\frac{-b'_i}{s_i}$ if $1 \leq i \leq r$, and $y_i =0$ otherwise, is feasible for \eqref{LP}
  with data $(S,b',c',0,0)$.
